Rough calculations

### 2015-10-31 notes/rough-calculations.Rmd
testing

*Useful functions used below:*
```{r setup_fns}
source("cline-fns.R")
```

@@

a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
b.  The selected cline takes time $s$ to spread out to its full width.
c.  Cline behaves like the selected cline for $r < 1/\tau$ (more tentatively, for $r < \sqrt{\frac{2}{s\tau}}$).
d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.

@@

d.  By L\'evy's identity, the local time of a Brownian motion at zero
    has the same distribution (pathwise!) as the running maximum of Brownian motion.
    Therefore, the local time at zero up until $\tau$ grows like $\sqrt{\tau}$.
    The same is true for Brownian motion reflected at zero.
    The time a lineage spends in heterozygotes should be roughly proportional to the local time
    (each time it visits zero from the left, it spends some time in hets before leaving again; 
    the constant of proportionality is the mean time spent per visit).
    Therefore, there will be appreciable chance of recombination if $r C \sqrt{\tau}>1$,
    where $C$ is the constant;
    this says that if clines down to $r$ are flat at $\tau$, then at time $2\tau$ they'll be flat down to $r/4$.

@@

>>>>

@@

a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
b.  The selected cline takes time $s$ to spread out to its full width.
c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2}{s\tau}}$ if $\tau>1/s$).
d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.

@@

d.  By Lévy's identity, the local time of a Brownian motion at zero
    has the same distribution (pathwise!) as the running maximum of Brownian motion.
    Therefore, the local time at zero up until $\tau$ grows like $\sqrt{\tau}$.
    The same is true for Brownian motion reflected at zero.
    The time a lineage spends in heterozygotes should be roughly proportional to the local time
    (each time it visits zero from the left, it spends some time in hets before leaving again; 
    the constant of proportionality is the mean time spent per visit).
    Therefore, there will be appreciable chance of recombination if $r C \sqrt{\tau}>1$,
    where $C$ is the constant;
    this says that if clines down to $r$ are flat at $\tau$, then at time $2\tau$ they'll be flat down to $r/4$.

@@

Simulation check
----------------

We'll check these with the PDE solutions
(and, then with simulations):

```{r setup_fns}
source("cline-fns.R")
```

a.  Look at space-time frequency images at $r=1$ and check they spread out as $\sqrt{t}$.

    ```{r unlinked_clines, cache=TRUE}
    thegrid <- setup.grid.1D(x.up=-50, x.down=50, N = 5000)
    s <- 0.01
    r <- 1.0
    tt <- seq(0,1000,length.out=250)
    unlinked.solns <- forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE)
    # image(unlinked.solns,legend=TRUE,zlim=c(0,1),yaxt='n')
    # axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
    unlinked.cline  <- cline_from_soln( unlinked.solns, grid=thegrid )
    image(unlinked.cline,legend=TRUE,zlim=c(0,1),yaxt='n',main=sprintf("r=%f, s=%f",r,s))
    axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
    matlines(tt,0.5+outer(sqrt(tt)/diff(range(thegrid$x.mid)),(1:4)/2,"*"),lwd=2,col='green')
    matlines(tt,0.5-outer(sqrt(tt)/diff(range(thegrid$x.mid)),(1:4)/2,"*"),lwd=2,col='green')
    ```

b.  Look at space-time frequency images at $r=0$ and several values of $s$
    to check they become stationary after about $1/s$.
    The clines are colored red for times less than $1/s$, and black afterwards;
    the green dotted lines are at $\pm 1/\sqrt{s}$.
    Indeed, cline flattening has slowed considerably by $1/s$, 
    and the limiting cline reaches $(1/4,3/4)$ at $\pm 1/\sqrt{s}$.

    ```{r forwards_selected_cline,cache=TRUE}
    thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
    svals <- exp(seq(log(1e-4),log(0.5),length.out=6))
    fwds.solns <- lapply(svals, function(s) {
            tt <- seq(0,3/s,length.out=40)
            forwards_pde(s=s,times=tt,grid=thegrid) 
        } )
    for (k in seq_along(svals)) {
        clineplot( fwds.solns[[k]], main=sprintf("s=%f",svals[k]), col=ifelse( fwds.solns[[k]][,1]>1/svals[k], adjustcolor("black",0.5), adjustcolor("red",0.5) ) )
        abline(v=c(-1,1)/sqrt(svals[k]),lty=3,col='green',lwd=2)
    }
    ```

c.  Look at space-recombination frequency images at a bunch of times
    to check that flattening occurs around $r=\sqrt{2/s\tau}$.
